---
title: "Data Arrangement ANSWER KEY"
author: "Chenxin Li"
date: "01/06/2023"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(tidyverse)
library(readxl)
```
## Data from lecture 
```{r}
child_mortality <- read_csv("../Data/child_mortality_0_5_year_olds_dying_per_1000_born.csv", col_types = cols()) 
babies_per_woman <- read_csv("../Data/children_per_woman_total_fertility.csv", col_types = cols()) 
```

These are two datasets downloaded from the [Gapminder foundation](https://www.gapminder.org/data/).
The Gapminder foundation has datasets on life expectancy, economy, education, and population across countries and years.
The goal is to remind us not only the "gaps" between developed and developing worlds, but also the amazing continuous improvements of quality of life through time.

1.  Child mortality (0 - 5 year old) dying per 1000 born.
2.  Births per woman.

These were recorded from year 1800 and projected all the way to 2100.

Let's look at them.

```{r}
head(child_mortality)
head(babies_per_woman)
```


```{r}
babies_per_woman_tidy <- babies_per_woman %>% 
  pivot_longer(names_to = "year", values_to = "birth", cols = c(2:302)) 

head(babies_per_woman_tidy)

child_mortality_tidy <- child_mortality %>% 
  pivot_longer(names_to = "year", values_to = "death_per_1000_born", cols = c(2:302)) 

head(child_mortality_tidy)
```

```{r}
birth_and_mortality <- babies_per_woman_tidy %>% 
  inner_join(child_mortality_tidy, by = c("country", "year"))

head(birth_and_mortality)
```

# Exercise

You have learned data arrangement! Let's do an exercise to practice what
you have learned today. 
As the example, this time we will use income per person dataset from Gapminder foundation.

```{r}
income <- read_csv("../Data/income_per_person_gdppercapita_ppp_inflation_adjusted.csv", col_types = cols()) 
head(income)
```

## Tidy data
Is this a tidy data frame?
NO! 

Make it a tidy data frame using this code chunk.
```{r}
income_tidy <- income %>% 
  pivot_longer(names_to = "year", values_to = "income", cols = !country)

head(income_tidy)
```

## Joining data

Combine the income data with birth per woman and child mortality data using this code chunk.
Name the new data frame "birth_and_mortality_and_income".

```{r}
 birth_and_mortality_and_income <- income_tidy %>% 
  inner_join(babies_per_woman_tidy, by = c("country", "year")) %>% 
  inner_join(child_mortality_tidy, by = c("country", "year"))

head(birth_and_mortality_and_income)
```
 

## Filtering data

Filter out the data for Bangladesh and Sweden, in years 1945 (when WWII ended) and 2010.
Name the new data frame BS_1945_2010.
How has income, birth per woman and child mortality rate changed during this 55-year period?

```{r}
BS_1945_2010 <- birth_and_mortality_and_income %>% 
 filter(country == "Bangladesh" | 
          country == "Sweden") %>% 
 filter(year == 1945 | 
         year == 2010)
 

head(BS_1945_2010)
```
 

## Mutate data

Let's say for countries with income between 1000 to 10,000 dollars per year, they are called "fed".
For countries with income above 10,000 dollars per year, they are called "wealthy".
Below 1000, they are called "poor".

Using this info to make a new column called "status".
Hint: you will have to use case_when() and the "&" logic somewhere in this chunk.

```{r}
birth_and_mortality_and_income <- birth_and_mortality_and_income %>% 
  mutate(status = case_when(
    income >= 1000 & income <= 10000 ~ "fed",       
    income > 10000 ~ "wealthy",                     
    income < 1000 ~ "poor"                        
))

head(birth_and_mortality_and_income)
```

## Summarise the data

Let's look at the average child mortality and its sd in year 2010. 
across countries across different status that we just defined. 
Name the new data frame "child_mortality_summmary_2010".

```{r}
child_mortality_summary_2010 <- birth_and_mortality_and_income %>% 
  filter(year == 2010) %>% 
  group_by(status) %>%
  summarize(
    avg = mean(death_per_1000_born), 
    sd = sd(death_per_1000_born))

head(child_mortality_summary_2010)
```
 
How does child mortality compare across income group in year 2010?
Child mortality is higher for lower income groups.
