---
title: "mean_separation_ANSWER_KEY"
author: "Chenxin Li"
date: "2023-01-07"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Packages 
```{r}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(ggbeeswarm)
```


# Data from lecture
```{r}
tissue_culture <- read_excel("../Data/LU-Matand - Stem Data.xlsx")
head(tissue_culture)
```

```{r}
tissue_culture %>% 
  mutate(Treatment = factor(Treatment, 
                            levels = c("T1", "T5", "T10"))) %>% 
  ggplot(aes(x = Treatment, y = Buds_Shoots)) +
  facet_wrap(~ Variety, scales = "free") +
  geom_quasirandom(aes(color = Explant), alpha = 0.8) +
  scale_colour_manual(values = brewer.pal(8, "Set2")) +
  labs(x = "Hormone Treatment",
       y = "Num buds and shoots",
       fill = NULL) +
  theme_classic() +
  theme(legend.position = "bottom") 

ggsave("../Results/04_multi_dots.png", height = 6, width = 9)
```
 

# Exercise
## Q1
Notice the `scales = "free` inside `facet_wrap()`. 
Make a new graph but remove the `scales = "free` inside `facet_wrap()`.


```{r}
tissue_culture %>% 
  mutate(Treatment = factor(Treatment, 
                            levels = c("T1", "T5", "T10"))) %>% 
  ggplot(aes(x = Treatment, y = Buds_Shoots)) +
  facet_wrap(~ Variety) +
  geom_quasirandom(aes(color = Explant), alpha = 0.8) +
  scale_colour_manual(values = brewer.pal(8, "Set2")) +
  labs(x = "Hormone Treatment",
       y = "Num buds and shoots",
       fill = NULL) +
  theme_classic() +
  theme(legend.position = "bottom") 

ggsave("Lesson4_answer1.png", height = 6, width = 9)
```

Can you explain what happened? 
Now all the varieties have the same y axis. 

Can you explain in what scenario is more appropriate to have `scales = "free"` turned on vs off? 
When we are only comparing within subplots, it's better to use `scales = "free"`. 
When we are comparing across subplots, it's better to not use `scales = "free"`.

## Q2 
Here is an example data for you to practice: 
```{r}
M1 <- data.frame(
  conc = rnorm(n = 8, mean = 0.03, sd = 0.01)
) %>% 
  mutate(group = "ctrl") %>% 
  rbind(
    data.frame(
      conc = rnorm(n = 6, mean = 0.25, sd = 0.02)
    ) %>% 
      mutate(group = "trt")
  ) %>% 
  mutate(pest = "Pest 1")

M2 <- data.frame(
  conc = rnorm(n = 8, mean = 6, sd = 1)
) %>% 
  mutate(group = "ctrl") %>% 
  rbind(
    data.frame(
      conc = rnorm(n = 6, mean = 5.5, sd = 1.1)
    ) %>% 
      mutate(group = "trt")
  ) %>% 
  mutate(pest = "Pest 2")

M3 <- data.frame(
  conc = rnorm(n = 8, mean = 20, sd = 0.5)
) %>% 
  mutate(group = "ctrl") %>% 
  rbind(
    data.frame(
      conc = rnorm(n = 6, mean = 19.5, sd = 1.2)
    ) %>% 
      mutate(group = "trt")
  ) %>% 
  mutate(pest = "Pest 3")

Spray <- rbind(
  M1, M2, M3
)

head(Spray)
```

In this hypothetical experiment, I have two treatments: ctrl vs. trt. 
And I measured the occurrence of three different pests after spraying either treatments. 
Make a mean separation plot for this experiment. 

Hints: 

* Is this a multifactorial experiment? Yes! 
* How many observations are there in each group? 

```{r}
Spray %>% 
  group_by(group, pest) %>% 
  count()
```
Ctrl groups has 8 observations, while trt group have 6 observations.  

Was the treatment effective in controlling any of the three pests? 
```{r}
Spray %>% 
  ggplot(aes(x = group, y = conc)) +
  facet_wrap(~ pest, scales = "free") +
  geom_quasirandom(aes(color = group), alpha = 0.8) +
  scale_colour_manual(values = brewer.pal(8, "Set2")) +
  labs(x = "Treatment",
       y = "concentration",
       fill = NULL) +
  theme_classic() +
  theme(legend.position = "bottom") 

ggsave("Lesson4_answer2.png", width = 4, height = 3)
```

Save the graph using `ggsave()`. 

 











