---
title: "Intro_to_heatmap_ANSWER_KEY"
author: "Chenxin Li"
date: "2023-01-15"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
# Packages 
```{r}
library(tidyverse)
library(RColorBrewer)
library(viridis) 
```

# Data from lecture 
```{r}
abc_1 <- expand.grid(
  a = c(1:10),
  b = c(1:10)
) %>% 
  mutate(c = a + b)

head(abc_1)
```

# Exercise 
## Q1
Here is the rainbow color scale: 
```{r}
abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = rainbow(20)) +
  labs(title = "c = a + b") +
  theme_classic() +
  coord_fixed()

ggsave("../Results/06_abc_2.png", height = 3, width = 3)
```
Can you explain why the rainbow color scale is inappropriate for heatmaps?  

There are multiple issues with the rainbow color scale.
1. It is not red/green colorblind friendly.
2. It is neither a unidirectional or bidirectional color scale. 

## Q2
You can ignore the following chunk. It generates the data. 
```{r}
Q2_data <- rbind(
  c(10, 20, 30, 40, 50, 60), # 6
  c(0, 10, 30, 10, 5, 0), # 3
  c(100, 200, 300, 400, 500, 550), # 6
  c(50, 45, 40, 30, 20, 0),  # 1
  c(0, 200, 300, 500, 200, 100), # 4
  c(10, 500, 200, 100, 10, 0), # 2 
  c(0, 0, 0, 10, 500, 0) # 5
) %>%
  as.data.frame() %>% 
  cbind(gene = c("a", "b", "c", "d", "e", "f", "g")) %>% 
  pivot_longer(cols = !gene, names_to = "V", values_to = "expression") %>% 
  mutate(stage = str_remove(V, "V"))

Q2_data %>% 
  ggplot(aes(x = stage, y = gene)) +
  geom_tile(aes(fill = expression)) +
  geom_text(aes(label = expression)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")))+
  theme_classic()

ggsave("../Results/06_genes_3.png", height = 3.5, width = 4.2)
```
What is (are) wrong with this heatmap? 
Hint: there is a lot going wrong with this heatmap. 

1. Values are not scaled to z-score. 
2. Bidirectional color scale used for unidirectional data.
3. Potential issue with outlier(s).  
