---
title: "Intro_to_network ANSWER KEY"
author: "Chenxin Li"
date: "2023-01-15"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages 
```{r}
library(tidyverse)
library(igraph)
library(ggraph)
library(readxl)
library(RColorBrewer)
```
# Homework 
## Q1 
Metabolic pathways can be modeled by networks, where each node is a metabolite and each edge is a metabolic enzyme. 
You can read more about how to represent metabolic pathways [here](https://github.com/cxli233/ggpathway). 

Here is an example:
```{r}
calvin_cycle_edges <- read_excel("../Data/Calvin_cycle_edges.xlsx")
calvin_cycle_nodes <- read_excel("../Data/Calvin_cycle_nodes.xlsx")

head(calvin_cycle_edges)
head(calvin_cycle_nodes)
```
Make a network diagram for the Calvin cycle. 
Color each node by how many carbons they have (the `carbon` column in the node table).
Label each metabolite using `geom_node_text(aes(label = name), hjust = 0.5, repel = T)`. 
Save the figure using `ggsave()`.
Use `bg = "white"` inside `ggsave()` to add a white background when exporting the png file.  

```{r}
network <- graph_from_data_frame(
  d = calvin_cycle_edges,
  vertices = calvin_cycle_nodes,
  directed = F)
```

```{r}
ggraph(network, layout = "kk") +
  geom_edge_link(color = "grey70", arrow = arrow(length = unit(3, "mm")))+
  geom_node_point(size = 3, shape = 21, color = "white", aes(fill = as.factor(carbon))) + 
  geom_node_text(aes(label = name), repel = T, hjust = 0.5) +
  scale_fill_manual(values = brewer.pal(8, "Set2")) +
  labs(fill = "carbon") +
  theme_void() 

ggsave("Lesson7_answer1.png", height = 4, width = 6, bg = "white")
```
 
## Q2 
[Phylogenetic trees](https://en.wikipedia.org/wiki/Phylogenetic_tree) and [dendrogram](https://en.wikipedia.org/wiki/Dendrogram) can be modeled by a network. 
Say we have a tree like this: 

```    
    I
    |
    G - H
    |
A - B - C 
    |   
    D - E
    |
    F
``` 

Make a network diagram and label each node.
Save the diagram using `ggave()`.
Use `bg = "white"` inside `ggsave()` to add a white background when exporting the png file.  
Hint: You can write the edge table in Excel and import it into R if that's easier. 
```{r}
edgetable <- data.frame(
  from = c("G", "G", "A", "B", "B","B", "D", "D"),
  to = c("I", "H", "B", "C", "G","D", "E", "F")
)
head(edgetable)
```

```{r}
nodes <- data.frame(
  nodes = unique(c(edgetable$from, edgetable$to))
)

nodes
```

```{r}
treenetwork <- graph_from_data_frame(
  d = edgetable,
  vertices = nodes,
  directed = F)
```

```{r}
ggraph(treenetwork, layout = "tree") +
  geom_edge_link(color = "grey70") +
  geom_node_point(size = 3, shape = 21, color = "white", fill = "grey20") +
  geom_node_text(repel = T, aes(label = nodes$nodes)) +
  theme_void()

ggsave("Lesson7_answer2.png", height = 3, width = 3.5, bg = "white")
```

