---
title: "08_Plot_assembly_ANSWER_KEY"
author: "Chenxin Li"
date: "2023-07-09"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 
```{r}
library(tidyverse)
library(ggbeeswarm)
library(patchwork)

library(RColorBrewer)
```

# Exercise
Use your own data from your research to generate a composite figure. 
Save it using `ggsave()`. 

For those of you who do not have your own data, please use these datasets from a hypothetical project.  

You measured tensile strength and resistance of a material at two temperatures (t1 and t2) and two pressures (p1 and p2) (dataset1). 
You also tested samples of this material in the above-mentioned temperature-pressure combinations. You recorded the percentages of samples that passed, failed, or were unclear (dataset2). 

* What is the condition that resulted in the highest tensile strength? 
* What is the condition that resulted in the lowest resistance? 
* Is there any correlation between tensile strength and resistance across various conditions? 
* What is the condition that resulted in the lowest percentage of failure? What is the breakdown of test outcomes across treatment? 
* Based on all the data available, which treatment is the best? 

## Datasets 
```{r}
dataset1 <- read.csv("../Data/dataset1.csv")
dataset2 <- read.csv("../Data/dataset2.csv")

head(dataset1)
head(dataset2)
```
### Panel A 
```{r}
panel_A <- dataset1 %>% 
  ggplot(aes(x = temperature, y = tensile_strength)) +
  facet_grid(.~ pressure) +
  ggbeeswarm::geom_quasirandom(alpha = 0.8, aes(color = treatment)) +
  scale_color_manual(values = brewer.pal(8, "Set2")) +
  labs(x = "Temperature",
       y = "Tensile strength") +
  theme_classic()

panel_A
```
### Panel B 
```{r}
panel_B <- dataset1 %>% 
  ggplot(aes(x = temperature, y = resistance)) +
  facet_grid(.~ pressure) +
  ggbeeswarm::geom_quasirandom(alpha = 0.8, aes(color = treatment)) +
  scale_color_manual(values = brewer.pal(8, "Set2")) +
  labs(x = "Temperature",
       y = "Resistance") +
  theme_classic()

panel_B
```
### Panel C
```{r}
panel_C <- dataset1 %>% 
  ggplot(aes(x = resistance, y = tensile_strength)) +
  facet_wrap(~ treatment) +
  geom_point(aes(color = treatment)) +
  labs(x = "Resistance",
       y = "Tesnsile strength") +
  scale_color_manual(values = brewer.pal(8, "Set2")) +
  theme_classic()

panel_C
```

### Panel D
```{r}
panel_D <- dataset2 %>% 
  pivot_longer(names_to = "treatment", values_to = "percent_samples", cols = !outcome) %>% 
  ggplot(aes(x = treatment, y = percent_samples)) +
  geom_bar(stat = "identity", position = "stack", aes(fill = outcome)) +
  scale_fill_manual(values = c("grey20", "tomato1", "dodgerblue4")) +
  labs(y = "Percent samples",
       x = "Treatment") + 
  theme_classic()

panel_D
```

### Assemble them
```{r}
 wrap_plots(
  panel_A + labs(tag = "A") + theme(legend.position = "none"),
  panel_B + labs(tag = "B") + theme(legend.position = "none"),
  panel_C + labs(tag = "C"),
  panel_D + labs(tag = "D"),
  guides = "collect",
  design = "AB
            CD")

ggsave("Lesson8_answer1.svg", height = 6, width = 8)
ggsave("Lesson8_answer1.png", height = 6, width = 8)
```

Based on all the data available, t2p2 appeared to the best performer out of all metrics studied. 
